---
title: "Case Study 2 Analysis"
author: "DW"
date: "2022-10-05"
output:   
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Case Study 2

Bellabeat provided smart device fitness data and requested insights on how consumers are using the smart devices.

Data was provided for the month of April 2016.

Data was processed and cleaned based on the following workflow:

```{r loading_packages, message=FALSE}
library(tidyverse)
library(dplyr)
library(tidyr)
library(skimr)
library(lubridate)
library(janitor)
library(dplyr)

# Extract CSV data
files = list.files(path = "C://Users//dlwan//Documents//Google_Certificate//Case2//Data")
setwd("C://Users//dlwan//Documents//Google_Certificate//Case2//Data")

file_use = list()
file_name = list()

# Check file list contains only csv
for (x in 1:length(files)) {
  ext = sapply(strsplit(files[x],split=".",fixed=TRUE),tail,1)
  name = sapply(strsplit(files[x],"_"),"[",1)
  if(ext == "csv") {
    file_use = append(file_use,files[x])
    file_name = append(file_name,name)
  }
}

id_cnt = list()
day_cnt = list()

time = list('ActivityDate','ActivityDay','Time','ActivityHour','ActivityMinute','SleepDay','Date')

# Store each csv to file name
# Convert Date to appropriate format
for (x in 1:length(file_use)) {
  temp = read_csv(file_use[x])
  
  # Remove duplicate values
  temp = distinct(temp)
  for (j in 1:length(time)){
    index = which(colnames(temp) == time[j])
    # Check if time index is in column names
    if (length(index)>0){
      # Add column for date in Date format
      # Remove NA
      test = paste(time[j])
      # print(sum(duplicated(temp)))
      if (grepl(":",temp[1,index])){
        temp = temp %>% 
          mutate(date_time = as.POSIXct(temp[[test]],format ="%m/%d/%Y %I:%M:%S %p", origin = '1970-01-01')) %>% 
          mutate(date_day = as.Date(date_time)) %>% 
          mutate(day_of_week = wday(date_day)) %>%
          
          na.omit()
      } else {
        temp = temp %>% 
          mutate(date_day = as.POSIXct(temp[[test]],format ="%m/%d/%Y", origin = '1970-01-01')) %>% 
          mutate(day_of_week = wday(date_day)) %>%
          na.omit()
      }
    } 
  }
  temp$Id = as.character(temp$Id)
  assign(unlist(file_name[x]),temp)
  id_cnt = append(id_cnt,n_distinct(get(unlist(file_name[x]))$"Id"))
  day_cnt = append(day_cnt,n_distinct(get(unlist(file_name[x]))$date_day))
}
```

## Daily Intensity 
Daily intensity data consisted of number of steps, distance, and calories.

Summary statistics for each member, including average number of steps, average distance, average calories, and number of recorded days.

Not every member used smart device each day.

```{r statistics_members, echo=FALSE}
sum_activity = dailyActivity %>% 
  group_by(Id) %>% 
  filter(TotalSteps>0) %>%
  summarize(Avg_Steps = mean(TotalSteps),Avg_Distance=mean(TotalDistance),Avg_Calories = mean(Calories),days=length(TotalSteps), Total_steps = sum(TotalSteps), Total_Distance=sum(TotalDistance), Total_Calories = sum(Calories),
            Logged_Distance = sum(VeryActiveDistance)+sum(ModeratelyActiveDistance)+sum(LightActiveDistance),
            V_dist_avg = sum(VeryActiveDistance)/Logged_Distance,
            m_dist_avg= sum(ModeratelyActiveDistance)/Logged_Distance,
            l_dist_avg= sum(LightActiveDistance)/Logged_Distance,
            Logged_Minutes = sum(VeryActiveMinutes)+sum(FairlyActiveMinutes)+sum(LightlyActiveMinutes),
            V_min_avg=sum(VeryActiveMinutes)/Logged_Minutes,
            m_min_avg=sum(FairlyActiveMinutes)/Logged_Minutes,
            l_min_avg=sum(LightlyActiveMinutes)/Logged_Minutes)
print(sum_activity[0:5], n=40)
```

Following visuals highlight key parameters for each member.

```{r daily_steps, echo=FALSE}
# Plot of Average number of steps vs Number of Active Days
ggplot(data=sum_activity)+
  geom_point(mapping=aes(x=days,y=Avg_Steps))+
  ggtitle("Average Steps vs Number of Active Days") + 
  xlab("Number of Active Days") +
  ylab("Average Number of Steps")
```

People who had an average number 6,000 steps a days were more likely to continuously use the smart device.


```{r monthly_calories, echo=FALSE}
# Plot of Total calories vs Number of Active Days
ggplot(data=sum_activity)+
  geom_point(mapping=aes(x=days,y=Total_Calories))+
  ggtitle("Total Calories vs Number of Active Days") + 
  xlab("Number of Active Days") +
  ylab("Total Calories")
```

People who had an average monthly 45,000 calories were more likely to continuously use the smart device.


```{r activity_intensity, echo=FALSE}
sum_intensity = dailyActivity %>% 
  group_by(Id) %>% 
  summarize(days=length(Id),
            Logged_Distance = sum(VeryActiveDistance)+sum(ModeratelyActiveDistance)+sum(LightActiveDistance)+sum(SedentaryActiveDistance),
            V_dist_avg = sum(VeryActiveDistance)/Logged_Distance,
            m_dist_avg= sum(ModeratelyActiveDistance)/Logged_Distance,
            l_dist_avg= sum(LightActiveDistance)/Logged_Distance,
            Logged_Minutes = sum(VeryActiveMinutes)+sum(FairlyActiveMinutes)+sum(LightlyActiveMinutes)+sum(SedentaryMinutes),
            V_min_avg=sum(VeryActiveMinutes)/Logged_Minutes,
            m_min_avg=sum(FairlyActiveMinutes)/Logged_Minutes,
            l_min_avg=sum(LightlyActiveMinutes)/Logged_Minutes,
            s_min_avg=sum(SedentaryMinutes)/Logged_Minutes)

## Plots of Intensity Distance Ratio vs Number of Active Days 
ggplot(data=sum_intensity)+
  geom_point(mapping=aes(x=days,y=V_dist_avg))+
  ggtitle("Intensity Distance vs Number of Active Days") + 
  xlab("Number of Active Days") +
  ylab("Proportion of Very Intense Distance")

## Plots of Intensity Time Ratio vs Number of Active Days 
ggplot(data=sum_intensity)+
  geom_point(mapping=aes(x=days,y=V_min_avg+m_min_avg))+
  ggtitle("Intensity Time vs Number of Active Days") + 
  xlab("Number of Active Days") +
  ylab("Proportion of Very Intense Time")
```  

Participants that spend 40 percent of the distance and  20 percent of active time at a high intensity would be more likely to use the smart device

## Conclusions
People who are more active in terms of further distance traveled or high degree of intensity are more likely to consistently use the smart device.
